
```{r}
library(tidyverse)
```

# 1.Obstacles to valid scientific inference
## Q1
1.Measurement distortions: vague standard
2.Selection bias: self select
3.Confounding variables: try to find cause and effect relationship, but there is another factor

# 2. paired t-test and effect size
```{r}
library(PairedData)
data("Barley")
detach('package:PairedData',unload=TRUE)
detach('package:MASS',unload=TRUE)
head(Barley,4)
```

## Q1
```{r}
t.test(x=Barley$Glabron, y=Barley$Velvet, paired = TRUE, conf.level = 0.99)
# 0.0118 > 0.01 H0 u = 0
```

## Q2
```{r}
Barley_diffs <- Barley %>%
  mutate(diffs=Glabron-Velvet) %>%
  pull(diffs)
y_bar <- mean(Barley_diffs)
s <- sd(Barley_diffs)
effect_size <- y_bar / s
effect_size
```

## Q3
```{r}
Barley <- Barley %>%
  mutate(diffs=Glabron-Velvet)

ggplot(data = Barley, aes(x = diffs)) + 
  geom_density() + 
  theme_bw() + 
  xlab("Difference in field")

ggplot(data = Barley, aes(sample = diffs)) + 
  theme_bw() +stat_qq()+ stat_qq_line(color="blue")
```

# 3. Implementing unpaired t-test

```{r}
library(palmerpenguins)
peng_AC<-penguins %>%
  drop_na(species,body_mass_g) %>%
  filter(species!="Gentoo")
head(peng_AC %>%select(species,flipper_length_mm,body_mass_g),5)
```

## Q1
```{r}
val_col<-"body_mass_g"
group_col<-"species"
data<-peng_AC

data_new<-data%>%
  rename(group=(!!group_col),val=(!!val_col))%>%
  group_by(group) %>%
  drop_na(val) %>%
  summarise(mn=mean(val))

data_new
```

```{r}
data_new$mn[2]
```

```{r}
t_test_function <- function(data, val_col, group_col) {
  data_new <- data %>% 
    rename(group=(!!group_col),val=(!!val_col))%>%
    group_by(group) %>%
    drop_na(val) %>%
    summarise(mn=mean(val), sd=sd(val), size=length(val)) 
  n_0 <- data_new$size[1]
  n_1 <- data_new$size[2]
  sd_0 <- data_new$sd[1]
  sd_1 <- data_new$sd[2]
  mean_0 <- data_new$mn[1]
  mean_1 <- data_new$mn[2]
  sd_combined <- sqrt(((n_0-1)*sd_0^2+(n_1-1)*sd_1^2)/(n_0+n_1-2))
  t_statistic <- (mean_0-mean_1)/(sd_combined*sqrt(1/n_0+1/n_1))
  p_value <- 2*(1-pt(abs(t_statistic),df=n_0+n_1-2))
  effect_size <- (mean_0-mean_1)/sd_combined
  final_result <- data.frame(t_stat=t_statistic, effect_size=effect_size, p_val=p_value)
  final_result
}
t_test_function(data=peng_AC,val_col="body_mass_g",group_col="species")
```

# 4.Useful concepts in statistical hypothesis testing

## Q2

(1) no  
(2) yes  

# 5.Investigating test size for an unpaired Student’s t-test

## Q1
```{r}
num_trials<-10000
sample_size<-30
mu_0<-1
mu_1<-1
sigma_0<-3
sigma_1<-3
alpha<-0.05
set.seed(0)

single_alpha_test_size_simulation_df<-data.frame(trial=seq(num_trials))%>%
  mutate(sample_0=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_0,sd=sigma_0)),sample_1=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_1,sd=sigma_1))) %>%
  mutate(p_value=pmap(.l=list(trial,sample_0,sample_1),.f=~t.test(..2,..3,var.equal=TRUE)$p.value))%>%
  mutate(type_1_error=p_value<alpha)

single_alpha_test_size_simulation_df %>%
  pull(type_1_error) %>%
  mean()
```

```{r}
num_trials <- 10000
sample_size <- 30
mu_0 <- 1
mu_1 <- 1
sigma_0 <- 3
sigma_1 <- 3
alpha_values <- seq(0.01, 0.1, 0.01)
set.seed(0)

compute_H0 <- function(alaph) {
  data.frame(trial = seq(num_trials)) %>%
    mutate(sample_0 = map(.x = trial, .f = ~ rnorm(n = sample_size, mean = mu_0, sd = sigma_0)),
    sample_1 = map(.x = trial, .f = ~ rnorm(n = sample_size, mean = mu_1, sd = sigma_1))) %>%
    mutate(p_value = pmap_dbl(.l = list(trial, sample_0, sample_1), 
                                .f = ~ t.test(..2, ..3, var.equal = TRUE)$p.value)) %>%
    summarise(type_1_error = mean(p_value < alaph)) %>%
    pull(type_1_error)
}

alpha_test_size_simulation_df <- data.frame(alpha = alpha_values) %>%
  mutate(type_1_error = map_dbl(.x = alpha, .f=compute_H0))

ggplot(alpha_test_size_simulation_df, aes(x = alpha, y = type_1_error)) +
  geom_line() +
  labs(x = "Significance Level (Alpha)", y = "Type I Error Rate") +
  theme_bw()

```

# 6.The statistical power of an unpaired t-test
