---
title: "Assignment 3"
author: "Xinwei Fan"
date: "2023-10-11"
output: html_document # you can change to other output format if you want
---


```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

# 1. Exploratory data analysis

```{r}
head(Hawks)
```
## 1.1 (Q1)   
<!-- Include your answers for each question under a second level heading eg ## 1.1 (Q1) -->
**** insert your solutions here *****
```{r}
HawksTail <- Hawks$Tail
HawksTail <- HawksTail[1:6]
HawksTail
tail_mean <- mean(HawksTail)
tail_mean
tail_med <- median(HawksTail)
tail_med
```
## 1.2 (Q1)
```{r}
Hawks %>%
  summarize(Wing_mean=mean(Wing, na.rm=TRUE), Wing_t_mean=mean(Wing, trim=0.50, na.rm=TRUE), Wind_med=median(Wing, na.rm=TRUE), Weight_mean=mean(Weight, na.rm=TRUE), Weight_t_mean=mean(Weight, trim=0.50, na.rm=TRUE), Weight_med=median(Weight, na.rm=TRUE))
```
## 1.2 (Q2)
```{r}
Hawks %>%
  group_by(Species) %>%
  summarize(Wing_mean=mean(Wing, na.rm=TRUE), Wing_t_mean=mean(Wing, trim=0.50, na.rm=TRUE), Wind_med=median(Wing, na.rm=TRUE), Weight_mean=mean(Weight, na.rm=TRUE), Weight_t_mean=mean(Weight, trim=0.50, na.rm=TRUE), Weight_med=median(Weight, na.rm=TRUE))
```
## 1.3 (Q1)
aA+b
```{r}
A <- mean(HawksTail)
result_1 <- A*2+3
result_1
HawksTail_y <- HawksTail*2+3
result_2 <- mean(HawksTail_y)
result_2
```
## 1.3 (Q2)
```{r}
tail_var<-var(HawksTail_y, na.rm=TRUE)
tail_var
tail_sd<-sd(HawksTail_y, na.rm=TRUE)
tail_sd
```
## 1.4
```{r}
hal<-Hawks$Hallux
hal<-hal[!is.na(hal)]
outlier_val<-100
num_outliers<-10
# 添加10个100的outlier
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
mean(hal)
mean(corrupted_hal)
num_outliers_vect<-seq(0,1000) 
means_vect<-c() 
for(num_outliers in num_outliers_vect) { 
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  means_vect<-c(means_vect,mean(corrupted_hal)) 
}
head(means_vect)
```
## 1.4 (Q1)
```{r}
medians_vect<-c() 
for(num_outliers in num_outliers_vect) { 
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  medians_vect<-c(medians_vect,median(corrupted_hal)) 
}
head(medians_vect)
```
## 1.4 (Q2)
```{r}
t_means_vect<-c() 
for(num_outliers in num_outliers_vect) { 
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  t_means_vect<-c(t_means_vect,mean(corrupted_hal, trim=0.1)) 
}
head(t_means_vect)
```
## 1.4 (Q3)
```{r}
df_means_medians<-data.frame(num_outliers=num_outliers_vect,mean=means_vect, t_mean=t_means_vect,median=medians_vect)
head(df_means_medians)
df_means_medians%>% pivot_longer(!num_outliers,names_to="Estimator",values_to="Value")%>% ggplot(aes(x=num_outliers,color=Estimator,linetype=Estimator,y=Value))+ geom_line()+xlab("Numberofoutliers")
```

median? because it's very steady

## 1.5 (Q1)
```{r}
ggplot(data=Hawks, aes(x=Species, y=Weight))+geom_boxplot()+xlab('Species') + ylab("Weight")
```

## 1.5 (Q2)
```{r}
Hawks %>%
  group_by(Species) %>%
  summarize(quantile025=quantile(Weight, probs = 0.25, na.rm=TRUE),
            quantile050=quantile(Weight, probs = 0.5, na.rm=TRUE),
            quantile075=quantile(Weight, probs = 0.75, na.rm=TRUE))
  
```
## 1.5 (Q3)
```{r}
num_outliers <- function(vec) {
  q25 <- quantile(vec, probs = 0.25, na.rm=TRUE)
  q75 <- quantile(vec, probs = 0.75, na.rm=TRUE)
  IQR <- q75 - q25
  outliers <- sum(vec < (q25 - 1.5 * IQR) | vec > (q75 + 1.5 * IQR), na.rm = TRUE)
  return (outliers)
}
num_outliers(c(0,40,60,185))
```
## 1.5 (Q4)
```{r}
Hawks %>%
  group_by(Species) %>%
  summarise(num_outliers_weight=num_outliers(Weight))
```
## 1.6 (Q1)
```{r}
S <- cov(Hawks$Weight, Hawks$Wing, use='complete.obs')
R <- cor(Hawks$Weight, Hawks$Wing, use='complete.obs')
S
R
```
## 1.6 (Q2)
acS
acR/|a||c|
```{r}
S_ <- cov(Hawks$Weight * 2.4 + 7.1, Hawks$Wing * -1 + 3, use='complete.obs')
R_ <- cor(Hawks$Weight * 2.4 + 7.1, Hawks$Wing * -1 + 3, use='complete.obs')
S_
R_
S_manual <- 2.4 * -1 * S
R_manual <- 2.4 * -1 * R / abs(2.4) * abs(-1)
S_manual
R_manual
```

# 2. Random experiments, events and sample spaces, and the set theory

## 2.1 (Q1)
random experiment: is a procedure  
1.has a well-defined set of possible outcomes;  
2.could (at least in principle) be repeated arbitrarily many times  
event: An event is a set (i.e. a collection) of possible outcomes of an experiment  
sample space: A sample space is the set of all possible outcomes of interest for a random experiment  
## 2.1 (Q2)
example: (1, 2)
sample space: 
{(1, 1), (1, 2), (1, 3), (1, 4), (1, 5), (1, 6),
(2, 1), (2, 2), (2, 3), (2, 4), (2, 5), (2, 6),
(3, 1), (3, 2), (3, 3), (3, 4), (3, 5), (3, 6),
(4, 1), (4, 2), (4, 3), (4, 4), (4, 5), (4, 6),
(5, 1), (5, 2), (5, 3), (5, 4), (5, 5), (5, 6),
(6, 1), (6, 2), (6, 3), (6, 4), (6, 5), (6, 6)}  
36  
no, because you can't roll empty.  

## 2.2 (Q1)  
1.
$\quad A \cup B = \{ 1,2,3,4,6 \}$  
$\quad A \cup C = \{1,2,3,4,5,6 \}$ 
2.  
$A \cap B= \{ 2 \}$
$A \cap C= \emptyset$
3.
$A\backslash B：\{1,3\}$
$A\backslash B：\{1,2,3\}$
4.
no
yes
5.
yes
6.
$\quad A \cup B$
$\quad A \cup B \cup C$  

## 2.2 (Q2)
1.A  
2.$\emptyset$  
3. let x in Bc, x is not B, A belongs to B, x is not A, so Bc belongs to Ac
# 3. Visualisation

## 3 (Q1)
```{r}
ggplot(data=Hawks, aes(x=Tail, color=Species))+
  geom_density()+theme_bw()+xlab("Tail (mm)")+ylab("Density")
```

## 3 (Q2)
```{r}
ggplot(data=Hawks, aes(x=Tail, y=Species, fill=Species))+geom_violin()+theme_bw()+xlab("Tail (mm)")+ylab("Density")
```

## 3 (Q3)
```{r}
mass_scatter <- ggplot(data=Hawks, aes(y=Weight, x=Tail))+
  xlab("Tail (mm)") + ylab("Weight (g)")

mass_scatter+geom_point(aes(color=Species, shape=Species))

```

## 3 (Q4)
```{r}
mass_scatter + geom_point(aes(color=Species)) + facet_wrap(~Species, scales = "free_y") + geom_smooth(method="lm")
```


## 3 (Q5)
```{r}
Hawks %>%
  select(Weight, Tail) %>% 
  filter(Weight==max(Weight, na.rm=TRUE)) 

mass_scatter <- ggplot(data=Hawks, aes(y=Weight, x=Tail))+
  xlab("Tail (mm)") + ylab("Weight (g)")

mass_scatter+geom_point(aes(color=Species)) + 
  geom_curve(x=196, xend=196, y=1800, yend=2030, arrow=arrow(length=unit(0.1, 'cm')), curvature=0.5) + 
  geom_text(x=196, y=1800, label="heaviest hawk")
```