
```{r}
library(tidyverse)
```

# 1.Maximum likelihood estimates
## 1.1 Maximum likelihood estimates for Redtailed hawks

```{r}
library(Stat2Data) 
data("Hawks")
```

## Q1

```{r}
RedTailedDf <- Hawks %>%
  filter(Species=="RT") %>%
  select(Weight, Tail, Wing)
head(RedTailedDf, 5)
```

## Q2

```{r}
n <- length(RedTailedDf$Tail)
mu_mle <- mean(RedTailedDf$Tail, na.rm = TRUE)
sigma_mle <-  sd(RedTailedDf$Tail, na.rm = TRUE) * sqrt((n-1)/n)
```

## Q3

```{r}
weights <- seq(mu_mle-3*sigma_mle, mu_mle+3*sigma_mle, 0.001*sigma_mle)

colors <- c("MLE density"="red", "Kernel density"="blue")

estimated_density <- data.frame(Length=weights, Density=dnorm(weights, mean = mu_mle, sd = sigma_mle))

plot_obj <- ggplot() + geom_line(data=estimated_density, aes(x=Length, y=Density, color="MLE density"))

plot_obj + geom_density(data=tibble(RedTailedDf$Tail), aes(x=RedTailedDf$Tail, color="Kernel density")) + labs(y="Tail length(mm)", color="colour") + theme_bw() + scale_color_manual(values=colors)
```

## 1.2 Unbiased estimation of the population variance

## Q1 

```{r}
set.seed(0)
trials <- 1000
sample_sizes <- seq(5, 100, 5)
mu <- 1
sigma <- 3

biasfunc <- function(x){return (x - sigma**2)}

df <- crossing(trial=seq(trials), sample_size=sample_sizes)
df <- mutate(df, simulation=pmap(.l=list(trial, sample_size), .f=~rnorm(.y, mean=mu, sd=sigma)))
sim_by_n_df <- df %>% 
  mutate(sample_var_vu = map_dbl(.x=simulation, ~var(.x))) %>%
  mutate(sample_var_mle = map2_dbl(.x=simulation, .y=sample_size, ~(sd(.x)*sqrt((.y-1)/.y))**2)) %>%
  mutate(bias_vu = map_dbl(.x = sample_var_vu, .f=biasfunc)) %>%
  mutate(bias_mle = map_dbl(.x = sample_var_mle, .f=biasfunc))

plot_obj_1_2_q1 <- ggplot() + labs(x="sample size", y="variance") + theme_bw() + geom_point(data=sim_by_n_df, aes(x=sample_size, y=bias_vu, color="Vu", linetype="Vu"), size=0.5) + geom_point(data=sim_by_n_df, aes(x=sample_size, y=bias_mle, color="Vmle", linetype="Vmle"),  size=0.5)

plot_obj_1_2_q1

```

## Q2

```{r}
set.seed(0)
trials <- 1000
sample_sizes <- seq(5, 100, 5)
mu <- 1
sigma <- 3

msefunc <- function(x){return (mean((x-sigma)**2))}
biasfunc <- function(x){return (x - sigma)}

df <- crossing(trial=seq(trials), sample_size=sample_sizes)
df <- mutate(df, simulation=pmap(.l=list(trial, sample_size), .f=~rnorm(.y, mean=mu, sd=sigma)))
sim_by_n_df <- df %>% 
  mutate(sample_var_vu = map_dbl(.x=simulation, ~sqrt(var(.x)))) %>%
  mutate(sample_var_mle = map2_dbl(.x=simulation, .y=sample_size, ~sd(.x)*sqrt((.y-1)/.y))) %>%
  mutate(bias_vu = map_dbl(.x = sample_var_vu, .f=biasfunc)) %>%
  mutate(bias_mle = map_dbl(.x = sample_var_mle, .f=biasfunc))

plot_obj_1_2_q2 <- ggplot() + labs(x="sample size", y="bias") + theme_bw() + geom_point(data=sim_by_n_df, aes(x=sample_size, y=bias_vu, color="Vu", linetype="Vu"), size=0.5) + geom_point(data=sim_by_n_df, aes(x=sample_size, y=bias_mle, color="Vmle", linetype="Vmle"),  size=0.5)

plot_obj_1_2_q2

```

## 1.3 Maximum likelihood estimation with the Poissond istribution

## Q3

```{r}
set.seed(0)
trials <- 1000
sample_sizes <- seq(5, 100, 5)
mu_0 <- 0.5

msefunc <- function(x){return (mean((x-mu_0)**2))}

df <- crossing(trial=seq(trials), sample_size=sample_sizes)
df <- mutate(df, simulation=pmap(.l=list(trial, sample_size), .f=~rpois(.y, mu_0)))
sim_by_n_df <- df %>% 
  mutate(sample_mean = map_dbl(.x=simulation, .f=mean)) %>%
  mutate(sample_mse = map_dbl(.x = sample_mean, .f=msefunc)) 

plot_obj_1_3_q3 <- ggplot() + labs(x="sample size", y="mse") + theme_bw() + geom_point(data=sim_by_n_df, aes(x=sample_size, y=sample_mse, color="sample_mse"), size=0.5)  

plot_obj_1_3_q3
```

## Q4

```{r}
data <- read.csv("VonBortkiewicz.csv")
fatalities <- data$fatalities
fatalities
lambda_MLE <- mean(fatalities)
lambda_MLE
prob_zero_fatalities <- dpois(0, lambda_MLE)
prob_zero_fatalities
```

## 1.4 Maximum likelihood estimation for the exponential distribution

## Q3
```{r}
data <- read.csv("CustomerPurchase.csv")
data <- data %>%
  mutate(time_diffs = lead(Time) - Time)
mu_mle <- length(data$time_diffs)/sum(data$time_diffs, na.rm = TRUE)
mu_mle
```

## Q4
```{r}
probability_exceed_one_minute <- 1 - pexp(60, rate = mu_mle)
probability_exceed_one_minute
```

# 2.Confidence intervals
## 2.1 Student’s t-confidence intervals

## Q2
```{r}
red_tailed_hawks_weights <- RedTailedDf$Weight[!is.na(RedTailedDf$Weight)]
ci <- t.test(red_tailed_hawks_weights, conf.level = 0.99)
ci
```

## Q3
```{r}
ggplot(data = RedTailedDf, aes(x = Weight)) + 
  geom_density() + 
  theme_bw() + 
  xlab("Weight of Red-Tailed Hawks")

ggplot(data = RedTailedDf, aes(sample = Weight)) + 
  theme_bw() +stat_qq()+ stat_qq_line(color="blue")
```

## 2.2 Investigating coverage for Student’s t intervals

```{r}
student_t_confidence_interval<-function(sample,confidence_level){ 
  sample<-sample[!is.na(sample)]
  n<-length(sample)
  mu_est<-mean(sample)
  sig_est<-sd(sample)
  alpha=1-confidence_level
  
  t<-qt(1-alpha/2,df=n-1)
  l=mu_est-(t/sqrt(n))*sig_est
  u=mu_est+(t/sqrt(n))*sig_est
  return(c(l,u))
}
```

## Q1
```{r}
num_trials<-100000
sample_size<-30
mu_0<-1
sigma_0<-3
alpha<-0.01
set.seed(0)

single_alpha_coverage_simulation_df<-data.frame(trial=seq(num_trials))%>%
  mutate(sample=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_0,sd=sigma_0)))%>%
  mutate(ci_interval=map(.x=sample,.f=~student_t_confidence_interval(.x,1-alpha)))%>%
  mutate(cover=map_lgl(.x=ci_interval,.f=~((min(.x)<=mu_0)&(max(.x)>=mu_0))))%>%
  mutate(ci_length=map_dbl(.x=ci_interval,.f=~(max(.x)-min(.x))))

single_alpha_coverage_simulation_df%>%
  pull(cover)%>%
  mean()
```

# 3.One sample hypothesis testing
## 3.1 One sample t-test on penguins data

## Q1

```{r}
library(palmerpenguins)
data("penguins")
```

```{r}
bill_adelie <- penguins %>%
  filter(species=="Adelie" & !is.na(bill_length_mm)) %>%
  select(bill_length_mm)

df_bill_adelie <- data.frame(bill_adelie)

ggplot(data = df_bill_adelie, aes(x = bill_length_mm)) + 
  geom_density() + 
  theme_bw() + 
  xlab("bill_length_mm")

ggplot(data = df_bill_adelie, aes(sample = bill_length_mm)) + 
  theme_bw() +stat_qq()+ stat_qq_line(color="blue")

ci <- t.test(bill_adelie, mu=40, conf.level = 0.99)
ci
```

## 3.2 Implementing a one-samplet-test

## Q1

```{r}
bill_adelie_values <- penguins %>%
  filter(species=="Adelie" & !is.na(bill_length_mm)) %>%
  pull(bill_length_mm)

compute_p_value <- function(sample, mu){
  sample_size <- length(sample)
  sample_mean <- mean(sample)
  sample_sd <- sd(sample)
  test_statistic <- (sample_mean - mu) / (sample_sd/sqrt(sample_size))
  p <- 2*(1-pt(abs(test_statistic),df=sample_size-1))
  return (p)
}
compute_p_value(bill_adelie_values, 40)
```
